---
title: "Regex Kurs"
author: "Leon Werner"
date: "25 11 2023"
output: html_document
---
content mostly based on the Text Analytics I HWS 22/23 excercise by Marlene Lutz
and partly on https://campus.datacamp.com/courses/intermediate-regular-expressions-in-r/regular-expressions-writing-custom-patterns?ex=2
# Background

Rgular Expressions (RegEx) allow to look for simple patterns in text.
E.g. 
- look for valid email adresses
- look for .de websites
- look for variable names in a big dataset


This is a markdown file, it is structured into executable code chunks and normal text (like this)

The next line is a code chunk but ignore this one for now
```{r setup, include=FALSE}
#this is onyl to set up the code, dont't touch this cell

#remove variables
rm(list = ls())

#load required packages, install only if not installed yet
p_needed = c(
  "tidyverse",
  "knittr",
  "devtools"
)
packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)] #all packages that are not installed yet

#install the uninstalled packages
if(length(p_to_install > 0)){
  install.packages(p_to_install)
}

#Install regexplain from github
if (FALSE == ("regexplain" %in% packages)){
  devtools::install_github("gadenbuie/regexplain")
}

#load all required packages
pkg_success <- sapply(p_needed, require, character.only = TRUE)

pkg_fail <- names(pkg_success)[which(!pkg_success)]

# set chunk options for the markdown file
knitr::opts_chunk$set(echo = TRUE)
```

# Recap of R
The course starts here
```{r recap R numers 1, include=True}
#this is a comment

this is code #but it doesn't work, put a hashtag before it to mark this line as comment!

# run code either line by line using ctrl + enter or chunk by chunk using the green play button
# you will see the results in the console

1 #this is value of 1 (also code). It will be "printed" (inclueded in the output)

print(1) #most of the times we don't have to write print

2+3 # use R as calculator, the result will be in the output
2*3
2**3

```

We don't really know which number belongs to which calculation. 
Maybe we should put some annotation in the output. This is where we need strings

```{r recap R strings 1, include=True}
"This is a string, a sequence of characters, it is encased by double quotation marks"
"Strings can include special characters !ยง$%&/()=? and numbers 1234567890"
"Strings can include most special characters like !ยง$%&/()=? and numbers 1234567890"

'It can also be included in single quotation marks'
"Or one insite the other 'like this'"
"Cannot use the same" "type of quotation marks twice" #if uncommented this line will cause an error
"we could discribe outputs, like: now we will print the number 1"
1

```

```{r recap R numers 2, include=True}
a = 5    # store values by assign them to "object" or "variable" a; variable names have to start with a letter
"value of a"
a        # get value of a
get("a") # get value of a
"value of 2*a"
2*a      # work with the value of the variable

a <- 2   #assign different value (using not = but <-)
"new value of a"
a
"sequence of numbers"
10:13
"a is a variable of class"
class(a)
```
```{r recap R lists, include=True}
list1 = c(a,2,a,4,5) #have a "list" or "vector" of values using c()
"value of list1"
list1
"new value of list1"
list1 = 1:5
list1
# select some values from the list unsing indices within []
"selected values from c"
list1[1]
list1[2:4]
"list1 is a variable of class"
class(list1)
"list1[1] is a variable of class"
class(list1[1])
```
Strings are also values. We can also assign them to variables

```{r recap R strings 2, include=True}

hello = "Hello" #assign string to variable 
world = "World!"
string_list = c(hello, world) # put string variables into one list
"List" 
string_list
helloworld = paste(hello, world) # we can also paste the strings together
"combinded (pasted) strings"
helloworld
"helloworld is a variable of class"
class(helloworld)
```
```{r recap R bool, include=True}
TRUE
FALSE #are the two boolean values
bool_list = c(TRUE,FALSE,FALSE,TRUE,FALSE)
list1[bool_list] #can use list of boolean values to choose from list (instead of index)
"TRUE is of class"
class(TRUE)
```

# Regex
The functions we use are from the tidyverse package

str_detect and grep check if a text contains a pattern
Try to figure out why we split the string first

```{r regex1, include=T}
a_txt = "This is an example text consisting fof 20 words with an average of 7 chars per word, test 1234123 1233 2312 1231 "

# we can split strings using str_split (e.g by pace " ")
"split output"
a_txt_split = str_split(a_txt," ")
a_txt_split #this is a nested list, lets change unlist it
"unlisted"
a_txt_split = unlist(a_txt_split)
a_txt_split

#we can use grep to get all elements in a list that fit with our re (or pattern)
# here we look for the occurence of "a" in the string
re = "a"
"outputs"
detect_output = str_detect(a_txt_split, pattern = re) #returns a list of Bools
detect_output
grep_output = grep(a_txt_split, pattern = re) #returns a list of indices
grep_output
#we can use both to select from the original list
"list selection"
a_txt_split[detect_output]
a_txt_split[grep_output]
```
The next methods stop after they have found something (and go to the next element in the list)
They all have verions that continue for all occurences of the pattern.
```{r start str_extract, include=T}
#This extracts whatever matched the pattern
str_extract(a_txt, pattern = re)
"version with all"
str_extract_all(a_txt, pattern = re)
```

```{r start str_replace, include=T}
#This replaces whatever matched the pattern
replacement = "" #replace with nothing (delete)
str_replace(a_txt, pattern = re, replacement = replacement)
str_replace(a_txt_split, pattern = re, replacement = replacement)
"version with all"
str_replace_all(a_txt, pattern = re, replacement = replacement)
str_replace_all(a_txt_split, pattern = re, replacement = replacement)

```

# Handy shiny-app
```{r regexplain, include=T}
regexplain::regexplain_gadget()
```

# Excercise 1
In this task you are asked to create regular expressions that meet the specified conditions.
__a)__ Write a regular expression that returns all integer numbers from a text that are surrounded by whitespaces. 
```{r t1a R, include=T}
a_txt = "This is an example text consisting fof 20 words with an average of 7 chars per word, test 1234123 1233 2312 1231 "

```

__b)__ Write a regular expression that returns all valid years that are surrounded by whitespaces in a text. A valid year is a 4 digit number in the range from 0000 to 2022. __(2 pts)__

```{r t1b R, include=T}
b_txt = "test 10001 0000 0100 0001 1111 0011 1234 1999 test 2000 test 2001 test 2010 test 2019 test 2022 test3 2023 test 2024 test test 9999 "

```

__c)__ Write a regular expression that returns all dates in the format YYYY-MM-DD or YYYY/MM/DD from a given text. Make sure that YYYY is a valid year (see task __b)__), MM is a valid month (1 to 12) and DD is a valid day (1 to 31). There is no need to make sure that e.g. XXXX-02-31 does not exist
```{r t1c R, include=T}
c_txt = "NOT VALID 12001-11-11 also not valid 2001-11-123 and not x2001-11-12-11 VALID 2022-12-31 2022/12/31 2022-09-31 2022-12-05 not valid 2023-12-31 2022-13-31 2022-12-32 1 2012/10-20 2012-10/20"

```


__d)__ Assume you are given a list ``l`` of strings like the one below. Using regular expressions, return a list that contains all elements from ``l`` that **don't contain both, the letter ``a`` AND ``e``** and store the result in a variable ``l_filtered``. 

 __Example:__ _given the list_

 ``l = ["apple", "cucumber", "tomato", "zucchini", "pumpkin", "pear", "raspberry", "blueberry"]``

 _you should return_

 ``l_filtered = ['cucumber', 'tomato', 'zucchini', 'pumpkin', 'blueberry']``.

# example list
l = ["apple", "cucumber", "tomato", "zucchini", "pumpkin", "pear", "raspberry", "blueberry"]

```{r t1d, include=T}
c_txt = "NOT VALID 12001-11-11 also not valid 2001-11-123 and not x2001-11-12-11 VALID 2022-12-31 2022/12/31 2022-09-31 2022-12-05 not valid 2023-12-31 2022-13-31 2022-12-32 1 2012/10-20 2012-10/20"

```


 __e)__ For the given string ``s`` with 4 lines, change the _whole_ word ``pot`` (i.e. ``pottery`` should not be changed) to ``1234`` only if it is at the start of a line. 
```{r t1e, include=T}

s = "
pottery clot pot
pot dot plot hot
spot rot pot got
not shot forgot"

```

#Excercise 2
```{r loading data, include=T}
#Load data
load("Thesis_data_changed.Rda")# Data from my thesis (text entries are in randomized order, demographic variables and edit time removed)

#column names names in data
colnames(df)

#we can use regex in select() and match()
IV03 = df %>% select( matches("IV03_01"))
df_without_IV03 = df %>% select( -matches("IV03_01"))
```

a) remove all TIME variables from df
```{r t2a, include=T}


```
b) create a second dataframe with only the first items from all variables
```{r t2b, include=T}


```

c) how many red dogs are mentioned in string_var_2 of df?
```{r t2c, include=T}


```

d) how many mentions of "Geld" or "Job" are in in IV03_01.
How many answers contain multiple mentions of "Job"
```{r t2d, include=T}


```

e) use ctrl+F to find all comments in this file that contain the string "regex"

#Excercise 3

See if a "FOUND" is followed by more than one "CHECK" before the next "FOUND" in the file Output.txt 

#Solutions
```{r Incomplete solutions Excercise 1, include=T}

#t1a = "(?<=\s)[0-9]+(?=\s)"
#t1b = "(?<=\s)([0-1]\d{3}|20[0-1]\d|202[0-2])(?=\s)"
#t1c = "(?:\b(?:[0-1]\d{3}|20[0-1]\d|202[0-2])-(?:0\d|1[12])-(?:0\d|[12]\d|3[01])\b)|(?:\b(?:[0-1]\d{3}|20[0-1]\d|202[0-2])/(?:0\d|1[12])/(?:0\d|[12]\d|3[01])\b)"
#t1c2 ="\b(?:[0-1]\d{3}|20[0-1]\d|202[0-2])[-/](?:0\d|1[12])(?:(?<=-..)-|(?<=/..)/)(?:0\d|[12]\d|3[01])\b")
# used ?: to include group seperator
# Using \b boundary to fix it to 4/2/2 pattern
# Backreference and [/-] is shorter but doesn't return a result in nice format

#t1d = "a.*e|e.*a" # a followed by e or e followed by a
#l_filtered = [w for w in l if not re_d.search(w)] 

#\b is word boundry
#t1e = "^pot\b", "1234" # this is wrong
```

